Robust Subspace and Higher Moments Distances for Botnet Detection

### 2019-04-21 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
adding texts of literature review

\title{Robust PCA and Higher Moments Distances for Botnet Detection}

@@

Some techniques can make the data discriminative for better anomaly detection: Data engineering (i.e. sampling \cite{??}, aggregation \cite{acarali2016survey}); Feature extraction (i.e. decomposition \cite{??}, robust estimates \cite{??}).

@@

@@

@@

@@

Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction. It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance. PCA has been used in attack detection \cite{almotairi2009technique}. However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.

Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.

Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack. Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.

@@

Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks. Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}. NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications. Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations. We propose to adopt CTU-13...

@@

\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)

@@

Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{moustafa2019holistic}.

@@

Based on related works \cite{lakhina2005mining,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.

>>>>

\title{Robust Subspace and Higher Moments Distances for Botnet Detection}

@@

Some techniques can make the data discriminative for better anomaly detection: Data engineering (i.e. sampling \cite{??}, aggregation \cite{callegari2011novel, acarali2016survey}); Feature extraction (i.e. decomposition \cite{??}, robust estimates \cite{??}).

@@

PCA is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}. However, PCA is not precise to reveal flows, time or detailed information for each anomlies, and sometimes relies into visual analysis for pricipal component selection.

@@

We believe that the distribution of anomalous and normal traffic are skewed but with differences that permit separate anomalous of normal traffic. (Assumption and hitotesis). We also believe that aggregated traffic and new features can make the data more discriminative for anomaly detection algorithms. We algo believe that normal traffic presents significant distance to anomalous traffic, which permit a distance based sepparation, that can be augmented by means of robust estimation of normal traffic, that can discard noise or irrelevant outliers and can highlight higher outliers when comparing to new observations

@@

Our approach train from normal traffic and have significant computational cost during training. For prediction, the cost is basically the cost for Mahalanobis distance computation

@@

Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction. It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance. PCA has been used in attack detection \cite{almotairi2009technique}. However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques. Callegari \emph{et al} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis. CAllegari separate normal of anomalies according to principal components (normal) and remaining (anomalies), based on PCA. It relies in the assumption that anomalies will cause a large change into anomalous components. This is similar to our previous work, but we rely on change into principal eigenvectors.

Use to justify the sensitivity of pca to outliers, only. Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack. Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.

@@

Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks. Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}. NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications. Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations. Considering that it is important to create intrusion detection datasets in modern-day computing to address the issues of DARPA/KDD. The next section discusses ne such contemporary dataset for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...

@@

\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).  -Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies. Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning. Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms. \textbf{To explain feature selection and aggregation}.

@@

Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}. We rely on this for our implementation

@@

Based on related works \cite{lakhina2005mining, callegari2011novel, chandrashekar2014survey, acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.
